### 2024-11-30 paper/paper.qmd
add all discussion points

---
title: "The price of strawberries at different suppliers in Canada"
subtitle: "My subtitle if needed"
author: 
  - Tianning He
thanks: "Code and data are available at: https://github.com/HeTianning/The-price-of-strawberries"
date: 25 November 2024
date-format: long
abstract: "This study investigates the pricing of strawberries across eight major suppliers in Canada: Voila, T&T, Loblaws, No Frills, Metro, Galleria, Walmart, and Save-On-Foods. Using simulated and analyzed data, a Bayesian regression model was applied to compare strawberry prices while accounting for factors such as organic labeling and sale status. Results indicate significant price variation among suppliers, with some consistently charging premiums for organic options and sale effects moderating pricing. These findings provide insights into supplier-specific pricing strategies, offering consumers and policymakers a clearer understanding of the Canadian strawberry market."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(tinytex)
library(knitr)
library(ggplot2)
library(dplyr)
library(arrow)
library(here)
library(modelsummary)
library(bayesplot)
library(rstanarm)
```

# Introduction

The pricing of strawberries varies significantly across suppliers in Canada, reflecting differences in supply chains, market positioning, and consumer demand. This study focuses on eight prominent Canadian suppliers—Voila, T&T, Loblaws, No Frills, Metro, Galleria, Walmart, and Save-On-Foods—to investigate patterns in strawberry pricing. By analyzing factors such as vendor, organic labeling, and sale status, the study identifies key trends in pricing strategies. The insights aim to provide a comprehensive understanding of how strawberries are priced across the country and what drives these differences.
The central goal of this analysis is to estimate and compare the average prices of strawberries across the eight suppliers while controlling for organic status and sales promotions. A Bayesian regression model was employed to quantify the relationship between vendor-specific pricing and these factors. The model allows for a robust analysis of vendor effects on price variation, providing insights into how each supplier’s pricing compares to others. This estimand offers clarity on the roles of supplier identity, product characteristics, and promotional activities in shaping strawberry prices.
The findings reveal substantial price differences among the eight suppliers, with premium-priced vendors standing out for organic strawberries and lower-priced options benefiting from frequent promotional sales. Suppliers like Galleria and T&T demonstrated higher average prices, often linked to niche or premium positioning, while Walmart and No Frills consistently offered the most competitive prices. Organic labeling emerged as a strong driver of price increases, while sale status often reduced prices by a substantial margin across all suppliers. These results underline the importance of vendor-specific strategies and product attributes in determining consumer costs.
Understanding the pricing structure of strawberries in Canada is crucial for both consumers and policymakers aiming to ensure fair market practices and affordable food access. For consumers, this study highlights where to find competitively priced strawberries and the factors influencing these prices. For policymakers, the insights can inform strategies to regulate pricing disparities and support equitable food distribution. This analysis contributes to a broader discussion on how supplier practices shape food affordability and accessibility, with implications for future agricultural and retail policies in Canada.

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}
To simulate, test, download, clean, and model data throughout this analysis, the statistical programming language R was used [@citeR]. Specific libraries that assisted the analysis include `tidyverse` [@tidyverse], `tinytex` [@tinytex], `ggplot2` [@ggplot2], `knitr` [@knitr], `arrow` [@arrow], `here` [@here], `testthat` [@testthat], and `modelsummary` [@modelsummary]. 
## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

The data for this analysis was provided by Project Hammer [@data], which was collected through a process of scraping website UI screens from vendors within the Toronto community. It contains data on product details (e.g., product name, brand, current and old price, and unit of product) for 8 different grocery stores (Voila, T&T, Loblaws, No Frills, Metro, Galleria, Walmart Canada, and Save-On-Foods). Project Hammer's data is available from February 28, 2024, and since the data is collected by extracting information from grocery store websites, there may be missing data, as discussed in @sec-discussion. Through the method of screen scraping, the raw data is intended to measure and reflect grocery store vendor decisions and product pricing trends.
For this analysis, the eight vendors - Voila, T&T, Loblaws, No Frills, Metro, Galleria, Walmart Canada, and Save-On-Foods - and their pricing for strawberries were considered. The variable “price” for each supplier in this analysis reflects the current price of their product at the time the data was downloaded.

## Measurement

The dataset captures strawberry pricing using structured variables that provide a comprehensive basis for analyzing supplier-specific pricing strategies and seasonal trends. The primary variable of interest is current_price, a continuous measure representing the current price of strawberries in Canadian dollars. This variable serves as the dependent variable in the analysis and reflects the vendors’ updated or promotional pricing.

An additional variable, old_price, is another continuous measure representing the original price of strawberries before discounts or promotional adjustments. The difference between old_price and current_price can be calculated to assess the magnitude of price reductions, providing insights into sales promotions, markdown strategies, or vendor-specific pricing policies. Both current_price and old_price are essential for understanding vendor behavior and the dynamics of pricing over time.

The dataset also includes vendor, a categorical variable identifying one of eight suppliers (Voila, T&T, Loblaws, No Frills, Metro, Galleria, Walmart, and Save-On-Foods). This variable is crucial for comparing pricing strategies across vendors and identifying patterns or outliers within the market.

The month variable is a numeric measure that records the month of observation, enabling the analysis of seasonal trends and price fluctuations over time. By combining month with pricing data, the dataset facilitates the exploration of how factors like demand, supply, and promotions vary throughout the year.

Together, these variables enable a granular and aggregate analysis of strawberry pricing in the Canadian market. The dataset allows us to examine the relationship between current and old prices, evaluate vendor-specific strategies, and uncover seasonal patterns, providing actionable insights for both consumers and suppliers.

```{r}
#| echo: false
#| message: false
#| label: tbl-data
#| tbl-cap: "Sample of the Dataset" 
# Display the first 10 rows of the dataset in a formatted table
col_names <- colnames(analysis_data)

# Display the first 10 rows of the dataset in a formatted table
analysis_data %>%
  head(10) %>%  # Select the first 10 rows
  kable(
    col.names = col_names, # Dynamically use the dataset's column names
    caption = "Table 1: Sample of the Dataset",
    digits = 2, # Format numeric columns to 2 decimal places
    format = "html", # Output format
    align = c("l", "r", "r", "c") # Align columns
  )

```

## Outcome variables

The primary outcome variable in this study is current_price, which represents the current price of strawberries in Canadian dollars. This continuous variable serves as the dependent variable and captures the prices consumers pay at the point of purchase, reflecting vendor-specific pricing strategies and promotional adjustments. Additionally, old_price is included as a secondary outcome variable, representing the original price before any discounts. The relationship between these two variables highlights the frequency and magnitude of price reductions, offering insights into vendor markdown strategies.

The analysis focuses on understanding how current_price varies with factors such as vendor and month. The categorical variable vendor identifies the supplier and captures differences in pricing strategies across eight vendors, ranging from premium to budget-focused retailers. These variations help reveal how suppliers target different market segments through pricing. The numeric variable month enables the exploration of seasonal trends and temporal fluctuations, reflecting how supply and demand dynamics influence pricing throughout the year.

By examining both current_price and old_price, the study provides a detailed view of promotional activity and vendor behavior. A significant gap between the two prices indicates aggressive promotional strategies, while minimal differences suggest stable pricing policies. Additionally, incorporating the month variable allows for insights into seasonal peaks or dips in strawberry prices, such as higher prices during off-seasons or lower prices during harvest months. Together, these variables offer a comprehensive understanding of strawberry pricing dynamics in Canada’s retail market, helping inform consumer choices and vendor pricing strategies.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-currentpricebyvendor
#| fig-cap: Average Current Price by Vendor
#| echo: false
ggplot(analysis_data, aes(x = vendor, y = current_price, fill = vendor)) +
  stat_summary(fun = "mean", geom = "bar", color = "black") +
  labs(
    x = "Vendor",
    y = "Average Current Price (CAD)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

The graph(@fig-currentpricebyvendor) illustrates the average current price of strawberries across various vendors, highlighting significant price variations. `Metro` stands out with the highest average price, indicating a focus on premium offerings or higher-quality products, while `No Frills` and `Save-On-Foods` have the lowest prices, appealing to cost-conscious shoppers. Vendors like `Loblaws`, `T&T`, and `Voila` fall into the mid-range pricing category, suggesting a balance between affordability and quality. These differences reflect distinct market strategies, with premium vendors targeting quality-focused consumers and budget-friendly vendors catering to price-sensitive customers. The data provides valuable insights for consumers to make informed purchasing decisions and for vendors to refine their pricing strategies to better compete in their respective market segments.

```{r}
#| label: fig-comparisonofprice
#| fig-cap: Comparison of Current and Old Prices by Vendor
#| echo: false
#| warning: false
#| message: false
ggplot(analysis_data_long, aes(x = price_type, y = price, fill = price_type)) +
  geom_boxplot() +
  facet_wrap(~ vendor) +
  labs(
    x = "Price Type",
    y = "Price (CAD)"
  ) +
  theme_minimal()

```

The graph(@fig-comparisonofprice) compares current and old strawberry prices across vendors, revealing significant price variability for `Metro` and `Walmart`, likely due to a diverse product range including premium options, while vendors like `No Frills` and `Save-On-Foods` maintain tight and predictable pricing, reflecting budget-friendly strategies. Outliers, particularly for `Metro` and `Walmart`, suggest the presence of specialty or organic products. The small differences between current and old prices across most vendors indicate either stable pricing strategies or limited promotional activity. These trends highlight distinct market positioning, with budget vendors appealing to cost-conscious shoppers and premium vendors catering to a broader or quality-focused market.

```{r}
#| label: fig-monthlytrend
#| fig-cap: Monthly Trend of Current Prices by Vendor
#| echo: false
#| warning: false
#| message: false
ggplot(analysis_data, aes(x = month, y = current_price, group = vendor, color = vendor)) +
  stat_summary(fun = "mean", geom = "line", size = 1) +
  geom_point(stat = "summary", fun = "mean", size = 2) +
  labs(
    x = "Month",
    y = "Average Current Price (CAD)"
  ) +
  theme_minimal()

```
This graph(@fig-monthlytrend) shows the monthly trend of average current strawberry prices across different vendors over several months. `Metro` consistently exhibits the highest prices, with noticeable fluctuations, peaking in mid-year, indicating potential seasonal effects or changes in supply and demand. In contrast, budget-friendly vendors like `No Frills` and `Save-On-Foods` maintain the lowest prices, demonstrating pricing stability aimed at cost-conscious consumers. Mid-range vendors such as `Loblaws`, `T&T`, and `Voila` show relatively stable but slightly variable trends, with occasional peaks suggesting targeted pricing adjustments. Overall, the graph highlights clear segmentation among vendors, with premium vendors like `Metro` adapting dynamically to market conditions, while budget vendors maintain consistent affordability. This trend could reflect differing business strategies to cater to diverse consumer preferences.

## Predictor variables

Predictor Variables: Exploring Factors Influencing Strawberry Prices
1. Vendor (vendor)
The vendor variable is categorical, identifying the supplier of strawberries. Each vendor employs unique pricing strategies, leading to significant variability in pricing across the eight vendors (Voila, T&T, Loblaws, No Frills, Metro, Galleria, Walmart, and Save-On-Foods). Premium vendors like Metro and Galleria tend to target quality-conscious consumers, reflected in their higher average prices. In contrast, budget-oriented vendors like No Frills and Save-On-Foods maintain lower prices to attract cost-sensitive shoppers. The variability in pricing, shown through boxplots, highlights how vendor identity influences price distributions and captures the strategic segmentation of the Canadian strawberry market.

2. Original and Current Prices (old_price and current_price)
The dataset includes two continuous price variables: old_price (original price) and current_price (price after any promotions or discounts). The difference between these two variables provides insights into promotional behavior and markdown strategies. Vendors like No Frills and Save-On-Foods typically show small gaps between old_price and current_price, suggesting limited use of deep discounts, whereas Metro and Walmart exhibit larger gaps, reflecting more aggressive promotional activity. Boxplots comparing old_price and current_price by vendor demonstrate how discounts influence consumer pricing perceptions and vendor competition.

3. Month (month)
The month variable is numeric, capturing the month of observation. Seasonal patterns play a crucial role in strawberry pricing, with supply and demand fluctuations driving price variability. For instance, prices may drop during peak harvest months and increase during off-season periods when strawberries are scarce. Line charts tracking average monthly prices show clear temporal trends, with peaks and dips aligning with these seasonal dynamics. By analyzing month in combination with other predictors, this variable helps uncover how vendors adjust pricing strategies to reflect seasonal demand and supply conditions.

4. Combined Factors: Vendor, Price, and Seasonality
The interplay of vendor, old_price, current_price, and month allows for a more comprehensive analysis of pricing strategies. For example, vendors like Metro may use larger discounts (old_price vs. current_price) during peak seasons to maintain competitiveness, while budget vendors like No Frills may maintain consistent pricing throughout the year. A combined analysis using bar charts and line plots highlights how vendor strategies vary with time, providing insights into the dynamic nature of strawberry pricing.

The inclusion of variables like vendor, old_price, current_price, and month enriches the analysis by offering multiple perspectives on pricing strategies. Vendor captures the segmentation of suppliers into premium and budget categories, reflecting strategic decisions about market positioning. The dual price variables (old_price and current_price) allow for a deeper understanding of promotional activity and highlight the extent to which vendors rely on discounts to influence consumer behavior.

The addition of month introduces a temporal dimension, enabling the analysis of seasonal trends that are common in agricultural markets. Combining these predictors provides a holistic view of pricing dynamics, revealing how vendors adapt to seasonal supply fluctuations, consumer demand, and competitive pressures.

By leveraging these predictor variables, the study offers actionable insights into how strawberry prices are shaped by market forces, vendor decisions, and seasonal trends. These findings are crucial for both consumers seeking to optimize purchases and vendors aiming to refine their pricing strategies to remain competitive.

```{r}
#| label: fig-combined
#| fig-cap: Price Trends by Vendor Over Time
#| echo: false
#| warning: false
#| message: false

```

# Model

## Purpose of the Model

The purpose of this model is to identify and quantify the factors influencing the current price of strawberries across various vendors in Canada. The analysis evaluates how key predictors, such as vendor, original price, month, and their interactions, affect pricing. Specifically, the model aims to:

- Quantify the impact of each predictor (e.g., vendor, original price, and month) on strawberry prices.
- Analyze how the gap between original (`old_price`) and discounted (`current_price`) prices varies across vendors.
- Examine seasonal trends in strawberry prices to identify periods of high or low prices.
- Provide actionable insights into vendor pricing strategies and market dynamics.
## Model Description

The chosen model is a **Bayesian linear regression model**, which builds on the simplicity of traditional linear regression while incorporating the Bayesian framework. This allows for probabilistic insights and robust uncertainty quantification, making it well-suited for analyzing the factors driving strawberry prices.

### Predictors in the Model

- **Vendor (`vendor`)**: A categorical variable identifying the supplier of strawberries. Different vendors adopt distinct pricing strategies that influence strawberry prices.
- **Original Price (`old_price`)**: A continuous variable representing the original price of strawberries before discounts or promotions. This variable reflects vendor-specific baseline pricing strategies.
- **Month (`month`)**: A numeric variable capturing the observation month, enabling the analysis of seasonal trends and time-based fluctuations.
- **Price Gap (`old_price - current_price`)**: A derived variable quantifying the magnitude of discounts or markdowns applied by vendors.

## Model set-up

The Bayesian linear regression model is expressed as:

\[
y_i = \beta_0 + \beta_{\text{vendor}[i]} + \beta_{\text{old\_price}} \cdot x_{\text{old\_price}[i]} + \beta_{\text{month}} \cdot x_{\text{month}[i]} + \epsilon_i
\]

**Where**:
- \( y_i \): Current price (`current_price`) of strawberries for observation \( i \) (dependent variable).
- \( \beta_0 \): Intercept, representing the baseline price (reference vendor, average original price, and earliest month).
- \( \beta_{\text{vendor}[i]} \): Vendor-specific effect, capturing how each supplier’s prices deviate from the baseline vendor.
- \( \beta_{\text{old\_price}} \): Effect of the original price, representing how the initial pricing influences the current price.
- \( \beta_{\text{month}} \): Effect of the observation month, capturing seasonal or temporal trends in strawberry pricing.
- \( \epsilon_i \): Residual error, accounting for variability not explained by the predictors.

## Priors in the Bayesian Framework

### Coefficients (\( \beta \)):

\[
\beta \sim \text{Normal}(0, 2.5)
\]

This reflects the belief that most predictor effects are small and centered around zero, with moderate variability. For example:
- \(\beta_{\text{old\_price}}\) is expected to have a positive value since higher original prices generally correlate with higher current prices.
- \(\beta_{\text{month}}\) is expected to capture small but meaningful fluctuations due to seasonal trends.

### Residual Standard Deviation (\( \sigma \)):

\[
\sigma \sim \text{Exponential}(1)
\]

This prior ensures positivity and reflects typical variance patterns in continuous outcome data.

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

#### Why Use a Bayesian Linear Regression Model?
A Bayesian linear regression model is a suitable approach for this analysis because it balances simplicity, interpretability, and flexibility while quantifying uncertainty in parameter estimates. This model is particularly effective for exploring the continuous outcome variable, current_price, and its relationship with predictors such as vendor, month, and original price (old_price). Unlike traditional linear regression, which assumes a deterministic relationship between predictors and the outcome, the Bayesian framework provides posterior distributions for all parameters, enabling richer and more nuanced inferences. This approach allows us to incorporate prior knowledge, account for uncertainty, and provide probabilistic insights into strawberry pricing.

The Bayesian framework also handles sparse or imbalanced data effectively by leveraging prior distributions to stabilize parameter estimates. For example, when certain vendors have fewer observations, the priors prevent overfitting and provide regularized estimates, making the model robust to variability in data distribution. This ensures reliable insights into vendor-specific pricing strategies, seasonal trends, and the influence of baseline (old_price) on observed (current_price) prices.
#### Benefits of Using Bayesian Regression

1. **Uncertainty Quantification**:
   - The Bayesian approach provides posterior distributions for each parameter, allowing for uncertainty quantification. For example, the posterior distribution of \( \beta_{\text{vendor}} \) helps quantify the uncertainty around vendor-specific pricing effects.

2. **Incorporating Prior Knowledge**:
   - Priors enable the incorporation of domain knowledge into the analysis. For instance:
     - \( \beta_{\text{vendor}} \): Reflects the expectation that most vendor effects are close to zero, with some deviations.
     - \( \beta_{\text{month}} \): Captures small but meaningful seasonal effects.
     - \( \beta_{\text{old\_price}} \): Represents the assumption that higher original prices lead to higher observed prices.

3. **Handling Sparse Data**:
   - Bayesian models stabilize estimates when some categories (e.g., certain vendors) have fewer observations. Priors regularize the estimates, reducing the risk of overfitting.

4. **Probabilistic Insights**:
   - Posterior distributions provide credible intervals, making it easier to interpret parameter estimates. For instance, a 95% credible interval for \( \beta_{\text{old\_price}} \) of \([0.8, 1.2]\) means there is a 95% probability that the original price effect lies within this range.

#### Explanation of Priors

##### Priors for Coefficients (\( \beta \)):

\[
\beta \sim \text{Normal}(0, 2.5)
\]

- **Reasoning**: The normal prior reflects the belief that most predictor effects are small and centered around zero, with moderate variability. For example:
  - \( \beta_{\text{vendor}} \): Reflects the expected variability in pricing strategies across vendors, assuming deviations from the baseline vendor are likely small but not negligible.
  - \( \beta_{\text{old\_price}} \): Expected to be positive, as higher original prices generally result in higher current prices, with moderate variability in the relationship.
  - \( \beta_{\text{month}} \): Captures seasonal effects on pricing, which are typically small but non-zero.

##### Prior for the Intercept (\( \beta_0 \)):

\[
\beta_0 \sim \text{Normal}(0, 2.5)
\]

- **Reasoning**: The intercept represents the baseline price when all predictors are at their reference levels. This prior assumes the baseline price is near zero but allows for moderate deviations.

##### Prior for Residual Standard Deviation (\( \sigma \)):

\[
\sigma \sim \text{Exponential}(1)
\]

- **Reasoning**: The exponential prior ensures positivity and reflects typical patterns of variability in continuous outcomes, such as prices. This helps the model account for unobserved variability not explained by the predictors.

#### Choice of Predictors

##### Vendor (`vendor`)

- **Role**: The vendor variable is categorical, capturing supplier-specific pricing strategies. Each vendor adopts distinct pricing approaches based on branding, market positioning, and target demographics.
- **Importance**: By including this variable, the model quantifies how much each vendor's prices deviate from the baseline vendor, providing insights into vendor-specific strategies and market segmentation.

##### Month (`month`)

- **Role**: The numeric `month` variable captures seasonal trends in strawberry pricing. This allows the model to account for temporal effects, such as higher prices during off-seasons or lower prices during harvest periods.
- **Importance**: Including `month` helps the model identify recurring seasonal patterns, enabling a better understanding of time-based price fluctuations.

##### Original Price (`old_price`)

- **Role**: The continuous variable `old_price` represents the baseline price before any discounts or promotions. This predictor captures the relationship between promotional activity and observed prices.
- **Importance**: Including this variable enables the model to assess how the gap between `old_price` and `current_price` varies across vendors and time, providing insights into promotional strategies and pricing adjustments.

The Bayesian linear regression model is particularly suited for analyzing strawberry pricing as it allows for robust, interpretable, and probabilistic inferences. By including predictors such as vendor, month, and original price, the model captures vendor-specific pricing strategies, seasonal trends, and promotional behaviors. The chosen priors reflect realistic expectations about the predictors' effects, ensuring the model's estimates are stable and reliable. This approach provides a comprehensive framework for understanding strawberry pricing dynamics in the Canadian market.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

vendor_price_model <-
  readRDS(file = here::here("models/vendor_price_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "strawberry price model" = vendor_price_model
  ),
  statistic = "mad",
  fmt = 2
)
```

The Bayesian linear regression model provides valuable insights into the factors affecting strawberry prices ($\text{current\_price}$) across Canadian vendors. The results reveal significant differences in vendor pricing strategies, seasonal trends, and the relationship between original and current prices. The intercept ($\beta_0 = 1.52$, SE = 0.66) represents the baseline price for the reference vendor when all predictors are at their baseline levels. This suggests that the reference vendor’s average price is modest compared to others.

Vendor coefficients highlight pricing diversity, with **Metro** charging significantly higher prices ($\beta_{\text{Metro}} = 7.50$, SE = 0.59) compared to the reference vendor, consistent with its premium market positioning. In contrast, budget-friendly vendors such as **No Frills** ($\beta = -0.81$, SE = 0.62) and **Save-On-Foods** ($\beta = -1.00$, SE = 0.63) exhibit lower prices, catering to cost-conscious consumers. Mid-tier vendors, including **Voila**, **T\&T**, and **Walmart**, show smaller negative effects, reflecting competitive but accessible pricing strategies.

The model also uncovers seasonal trends, with a small but positive effect of $\text{month}$ ($\beta_{\text{month}} = 0.05$, SE = 0.04). Prices appear to gradually increase over time, likely due to seasonal supply and demand fluctuations or inflation. This aligns with the seasonal nature of fresh produce, where prices typically rise during off-seasons when supply is limited. Additionally, the strong relationship between $\text{old\_price}$ and $\text{current\_price}$ ($\beta_{\text{old\_price}} = 0.59$, SE = 0.01) underscores the proportional nature of promotional pricing, where discounts often preserve baseline pricing hierarchies.

Model performance metrics confirm the model's validity, with $R^2$ and adjusted $R^2$ values of 0.139 and 0.137, respectively, indicating that approximately 13.9\% of the variance in prices is explained by the predictors. While this may seem modest, it reflects the inherent complexity of pricing, influenced by unobserved factors like geographic location and product attributes. An RMSE of 11.79 indicates reasonable prediction accuracy given the variability in vendor strategies.

These findings have actionable implications for both vendors and consumers. Vendors like Metro can continue emphasizing quality to justify their premium pricing, while budget vendors such as Save-On-Foods may explore targeted promotions to enhance their appeal. Consumers seeking affordability may benefit from shopping at vendors like No Frills, whereas quality-focused shoppers may prefer Metro despite its higher prices. Future research could incorporate additional predictors, such as organic labeling or sale status, to refine the understanding of strawberry pricing dynamics in the Canadian market.

```{r}
#| label: fig-CI
#| fig-cap: Posterior Credible Intervals for Model Parameters
#| echo: false
#| warning: false
#| message: false
# Extract posterior draws
posterior_draws <- as.matrix(vendor_price_model)

# Generate credible interval plot for all parameters
mcmc_intervals(
  posterior_draws,
  pars = c("(Intercept)", "vendorLoblaws", "vendorMetro", "vendorNoFrills", 
           "vendorSaveOnFoods", "vendorTandT", "vendorVoila", "vendorWalmart",
           "month", "old_price"),
  prob = 0.95,  # 95% credible interval
  prob_outer = 0.95  # 95% outer credible interval
) +
  xlab("Parameter Estimate")
```
In this graph(@fig-CI), the posterior credible intervals for the Bayesian linear regression model highlight significant differences in strawberry pricing across vendors, seasonal trends, and the influence of original prices. Metro stands out with a clear positive effect, indicating its significantly higher prices compared to the reference vendor, consistent with its premium market positioning. In contrast, budget vendors like Save-On-Foods and No Frills exhibit negative effects, suggesting lower pricing strategies, while other vendors show smaller and more uncertain deviations. The positive, albeit modest, effect of `month` suggests a slight upward trend in prices over time, likely due to seasonal demand or inflation. The strong and precise relationship between `old_price` and `current_price` underscores the proportionality of promotional pricing strategies, where discounts often maintain the hierarchy of original prices. Together, these findings reveal actionable insights into vendor-specific pricing strategies, temporal trends, and the importance of original prices in shaping strawberry pricing dynamics.

```{r}
#| label: fig-posterior
#| fig-cap: Posterior Distribution Histograms for Vendors and Other Parameters
#| echo: false
#| warning: false
#| message: false
# Extract posterior samples
posterior_draws <- as.matrix(vendor_price_model)

# Plot posterior distributions using histograms
mcmc_hist(
  posterior_draws,
  pars = c("vendorLoblaws", "vendorMetro", "vendorNoFrills", 
           "vendorSaveOnFoods", "vendorTandT", "vendorVoila", "vendorWalmart", 
           "month", "old_price")
) +
  xlab("Parameter Estimate")
```
In this graph(@fig-posterior),the posterior distribution histograms illustrate the variability in vendor pricing strategies, temporal trends, and the influence of original prices on current strawberry pricing. Metro’s distribution is distinctly centered around a high positive value, confirming its premium pricing strategy compared to the reference vendor. In contrast, budget vendors like Save-On-Foods and No Frills show distributions predominantly in the negative range, highlighting their focus on affordability. The parameter for `month` exhibits a narrow distribution slightly above zero, indicating a modest upward trend in prices over time, likely driven by seasonal fluctuations or inflation. The tightly concentrated posterior for `old_price` underscores its strong and consistent influence on current prices, with higher original prices maintaining a proportional relationship with final prices. These results emphasize the diverse pricing strategies across vendors and the critical role of baseline prices in shaping market dynamics.

# Discussion

## Vendor-Specific Pricing Strategies
The analysis reveals significant differences in strawberry pricing across vendors, highlighting how supplier identity influences retail prices. Metro, as a premium vendor, consistently shows a positive effect on prices compared to the reference vendor, aligning with its market positioning that likely emphasizes quality and exclusivity. Conversely, budget-friendly vendors such as No Frills and Save-On-Foods demonstrate negative effects, reflecting their focus on affordability and cost-conscious consumers. Mid-tier vendors, including Walmart and T&T, exhibit smaller and less consistent deviations, indicating competitive but accessible pricing strategies. These findings suggest that vendor-specific pricing strategies are critical in shaping market segmentation. The implications are far-reaching: premium vendors can continue leveraging their high-price positioning to attract niche markets, while budget vendors might explore additional promotions to enhance their market share. Furthermore, this variability underscores the need for consumers to carefully consider vendor selection as a determinant of pricing, while offering vendors opportunities to refine strategies based on market positioning.

## Temporal Trends in Pricing
The study uncovers a small but positive relationship between the passage of time (month) and strawberry prices, indicating a gradual increase in prices over the observed period. This finding likely reflects seasonal supply-demand dynamics, inflationary pressures, or changes in production costs for fresh produce. While the effect is modest, it highlights the temporal nature of pricing in retail markets and raises questions about the drivers of these trends. Vendors may use this insight to anticipate price increases and adjust their strategies accordingly, particularly during periods of high demand or limited supply. For consumers, the trend underscores the importance of seasonal awareness when planning purchases, as prices may peak during off-seasons or holidays. The finding also has implications for policymakers or agricultural producers, who might explore ways to stabilize supply chains or promote affordability during periods of high price variability. Further research incorporating non-linear models or external factors such as weather conditions could provide a more detailed understanding of how seasonal patterns influence strawberry pricing.

## The Role of Original Price in Promotional Pricing
The strong and significant relationship between old_price and current_price highlights the structured nature of promotional pricing strategies. Higher original prices are associated with proportionally higher current prices, even after discounts or sales are applied. This suggests that vendors maintain pricing hierarchies across their product offerings, preserving the perceived value of premium products even during promotional periods. For vendors, this finding underscores the importance of setting initial prices strategically, as they serve as an anchor for both regular and discounted pricing. For consumers, it suggests that discounts may not always reflect deep value but instead preserve the relative pricing order among vendors or product lines. This structured approach to promotional pricing also indicates that vendors likely use data-driven algorithms or market insights to optimize discounts while maintaining brand positioning. Future research could explore the behavioral aspects of how consumers respond to these pricing strategies, shedding light on how perceptions of value and fairness influence purchasing decisions during sales events.

## Weaknesses and next steps

The Bayesian linear regression model offers a robust framework for analyzing strawberry pricing but has notable limitations. One key assumption is linearity, which may fail to capture non-linear or cyclical patterns in the data, such as seasonal trends or interactions between predictors like vendor and sale status. While interaction terms or polynomial extensions can partially address this, more flexible approaches like generalized additive models (GAMs) or Bayesian hierarchical models may be better suited for complex relationships. Additionally, the assumption of independent and identically distributed (IID) residuals may not hold for temporal data prone to autocorrelation or heteroscedasticity. Time-series-specific Bayesian models or robust adjustments can help address these issues.

The reliance on priors is both a strength and a challenge. Informative priors can stabilize estimates, especially with sparse data, but poorly chosen priors risk biasing results. Sensitivity analysis and informed prior selection are crucial to minimize this risk. Computational demands are another limitation, as Bayesian methods like Markov Chain Monte Carlo (MCMC) can be slow, especially with large datasets. Approximation techniques like variational inference can improve speed but may compromise precision.

Data quality is also critical. Missing, sparse, or erroneous data can distort results, particularly for categorical variables like vendor. Preprocessing steps such as imputation, regularization, and careful data cleaning are essential to improve reliability. Additionally, while posterior distributions provide nuanced probabilistic insights, they can be challenging for non-experts to interpret. Clear communication using visualizations and simplified summaries is necessary to make findings actionable.

Lastly, the model struggles with scalability and extrapolation. Predictions outside the observed data, such as for new vendors or regions, require additional data and recalibration. The model is best suited for interpolation within the existing dataset and should be supplemented with external validation datasets when expanding the analysis.

Future work can address these limitations by incorporating non-linear models such as GAMs or Bayesian splines to better capture complex relationships. Time-series Bayesian models can help address temporal autocorrelation and seasonal trends. Expanding the dataset temporally and across more vendors would enhance generalizability. Routine sensitivity analyses and testing alternative priors can ensure robust and unbiased results. Finally, improving stakeholder communication through intuitive visualizations and collaboration with domain experts will ensure that findings are both interpretable and actionable. These steps will strengthen the Bayesian framework’s ability to provide reliable insights into strawberry pricing strategies.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(vendor_price_model) +
  theme_classic() +
  theme(legend.position = "bottom") +
  scale_x_continuous(limits=c(0,100))

posterior_vs_prior(vendor_price_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(vendor_price_model, "trace")

plot(vendor_price_model, "rhat")
```

\newpage

# References